---
title: "Sleep"
author: "Zihao Yu"
format: html
editor: visual
---

# Approach

## Sleep efficiency dataset

I've chosen this dataset because a lot of people stay up late. By looking at the data, it records the time spent in each sleep stage, so then I want to see how different factors(caffeine, alcohol consumption, smoking status, and exercise) affect sleep quality.

## Compare the effects of various factors on sleep states

Sleep dataset: https://raw.githubusercontent.com/XxY-coder/data607-week1/refs/heads/main/Sleep_Efficiency.csv

#view the columns and delete some columns.

```{r}
library(tidyverse)
library(gt)

url <- "https://raw.githubusercontent.com/XxY-coder/data607-week1/refs/heads/main/Sleep_Efficiency.csv"
df <- read_csv(
  file = url
)

glimpse(df)
```

##Cleaning the data
Caffeine, alcohol, smoking, and exercise are more straight forward, then rename them for convenient.
check and remove the NA values.


```{r}
df2 <-
  df |>
  select(
    Age,
    Gender,
    sleep_time = 'Sleep duration',
    caffeine_intake = 'Caffeine consumption',
    alcohol_intake  = 'Alcohol consumption',
    smoking_status  = 'Smoking status' ,
    exercise_freq   = 'Exercise frequency'
) |>
  filter(
    !is.na(caffeine_intake),
    !is.na(alcohol_intake),
    !is.na(smoking_status),
    !is.na(exercise_freq)
  )

glimpse(df)
```

####Check for relevance
Then compare how each factors effact to the sleep time. 
```{r}
ggplot(df2, aes(Age, sleep_time)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  labs(title = "sleep_time vs Age", x = "Age", y = "sleep_time")

ggplot(df2, aes(caffeine_intake, sleep_time)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm" ) +
  labs(title = "sleep_time vs Caffeine", x = "caffeine_intake", y = "sleep_time")

ggplot(df2, aes(alcohol_intake, sleep_time)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm" ) +
  labs(title = "sleep_time vs alcohol_intake", x = "alcohol_intake", y = "sleep_time")

ggplot(df2, aes(factor(exercise_freq), sleep_time, fill = exercise_freq)) +
  geom_boxplot() +
  labs(title = "sleep_time vs exercise_freq", x = "exercise_freq", y = "sleep_time")
df2 |>
  count(exercise_freq)
```

#####Conclusions
Four sets of comparisons reveal that overall correlations between age, caffeine intake, smoking status, and sleep duration are relatively weak.
The scatter plot shows no discernible trend, with only slight variations in data distribution. 
Box plots for exercise frequency indicate similar ranges across groups, with most sleep durations concentrated within the 7–8 hour range.
